Making ice sheet models \\ scale properly

### 2023-01-06 talk/slides.tex
a bit more

\newcommand{\bF}{\mathbf{F}}
\newcommand{\bV}{\mathbf{V}}
\newcommand{\bX}{\mathbf{X}}

@@

\bigskip
\item<2> {\footnotesize for more on this free-boundary perspective, see Jouvet \& Bueler (2012) and Bueler (2021)}
\end{itemize}
\end{frame}

@@

@@

\begin{frame}{complexity table}

%\setlength{\tabcolsep}{5pt}
%\renewcommand{\arraystretch}{1.5}

\begin{tabular}{llll}
\emph{time-stepping} & \emph{dynamics} & \emph{flops per model year} & \emph{[pessimistic stability]} \\ \hline
explicit & SIA    & $\oo{\frac{D\, L^2}{\Delta x^4}} = \oo{\frac{D\, m^2}{L^2}}$ \\
explicit & Stokes & $\oo{\frac{U L^{2+2\alpha}}{\Delta x^{3+2\alpha}}} = \oo{\frac{U m^{1.5+\alpha}}{L}}$ & $\oo{\frac{D\, L^{2+2\alpha}}{\Delta x^{4+2\alpha}}} = \oo{\frac{D\,m^{2+\alpha}}{L^2}}$ \\
implicit & SIA    & $\oo{\frac{q\, L^{2+2\beta}}{\Delta x^{2+2\beta}}} = \oo{q\, m^{1+\beta}}$ \\
implicit & Stokes & $\oo{\frac{q\, L^{2+2\gamma}}{\Delta x^{2+2\gamma}}} = \oo{q\, m^{1+\gamma}}$
\end{tabular}

\begin{itemize}
\item Asymptotic estimates of algorithmic scaling, measured by floating point operations per model year, for map-plane (2D) time-stepping numerical ice sheet simulations, in the high resolution limit where $\Delta x\to 0$ and $m\to\infty$
\end{itemize}
\end{frame}

@@

>>>>

\newcommand{\bF}{\mathbf{F}}
\newcommand{\bU}{\mathbf{U}}
\newcommand{\bV}{\mathbf{V}}
\newcommand{\bX}{\mathbf{X}}

@@

\bigskip
\item<2> {\footnotesize for more on this free-boundary perspective see Bueler (2021)}
\end{itemize}
\end{frame}

@@

\begin{itemize}
\item a helpful viewpoint on this complicated system is the equation

FIXME

where $\bU$ is vertically-averaged horizontal velocity within icy domain $\Lambda(t)$
\item \emph{question:} is it really an advection?

FIXME

\item[] \emph{answer:} no, because ice flows (mostly) downhill

\item alternative equation, the \emph{shallow diffusion} model of ice sheets:

FIXME
\bigskip
\item<2> {\footnotesize for more on the shallow diffusion perspective see Jouvet \& Bueler (2012)}
\end{itemize}
\end{frame}

\begin{frame}{regarding PDE solver complexity}

@@

\begin{frame}{complexity of ice sheet models}

\begin{tabular}{llll}
\emph{time-stepping} & \emph{dynamics} & \emph{flops per model year} \\ \hline
\onslide<1->{
\,explicit & SIA    & $\oo{\frac{D\, L^2}{\Delta x^4}} = \oo{\frac{D\, m^2}{L^2}}$ \\
}
\onslide<1->{
explicit & Stokes (\emph{optimistic}) & $\oo{\frac{U L^{2+2\alpha}}{\Delta x^{3+2\alpha}}} = \oo{\frac{U m^{1.5+\alpha}}{L}}$ \\
}
\onslide<2->{
explicit & Stokes (\emph{pessimistic})  & $\oo{\frac{D\, L^{2+2\alpha}}{\Delta x^{4+2\alpha}}} = \oo{\frac{D\,m^{2+\alpha}}{L^2}}$ \\
}
\onslide<3->{
implicit & SIA    & $\oo{\frac{q\, L^{2+2\beta}}{\Delta x^{2+2\beta}}} = \oo{q\, m^{1+\beta}}$ \\
}
\onslide<3->{
implicit & Stokes & $\oo{\frac{q\, L^{2+2\gamma}}{\Delta x^{2+2\gamma}}} = \oo{q\, m^{1+\gamma}}$
}
\end{tabular}

FIXME asymptotic estimates of algorithmic scaling, measured by floating point operations per model year, for map-plane (2D) time-stepping numerical ice sheet simulations, in the high resolution limit where $\Delta x\to 0$ and $m\to\infty$
\end{frame}

@@

\begin{frame}{obstacle problem solver scaling}
